---
title: Coursera Dec 2014  Practical Machine Learning
date: "Sunday, December 21, 2014"
output: html_document
---

Both the test and training sets are initially read into r:

```{r}
library(randomForest)

test<-read.csv("pml-testing.csv")
train<-read.csv("pml-training.csv")

dim(train)
```

First some data tidying.

The first 7 columns are not quantitative data, therefore excluded.  Likewise
all 'calculated' data (ie avg, kurtosis, max, min) is removed.  Just the 'raw' data remains:
```{r}
traindf1<-train[8:160]
traindf2<-traindf1[,-grep(pattern='avg',x=colnames(traindf1))]
traindf3<-traindf2[,-grep(pattern='max',x=colnames(traindf2))]
traindf4<-traindf3[,-grep(pattern='min',x=colnames(traindf3))]
traindf5<-traindf4[,-grep(pattern='stddev',x=colnames(traindf4))]
traindf6<-traindf5[,-grep(pattern='skewness',x=colnames(traindf5))]
traindf7<-traindf6[,-grep(pattern='kurtosis',x=colnames(traindf6))]
traindf8<-traindf7[,-grep(pattern='accel',x=colnames(traindf7))]
traindf9<-traindf8[,-grep(pattern='var',x=colnames(traindf8))]
traindf10<-traindf9[,-grep(pattern='amplitude',x=colnames(traindf9))]
dim(traindf10)
```

There are now 36 predictors and 1 outcome (classe).  Due to the large number of predictors, a random forest approach is selected.  

```{r}
fit<-randomForest(classe~.,data=traindf10,ntree=200,keep.forest=TRUE)
```

A fit is obtained, with an estimated out-of-bag error rate of 0.32%.  Now, the test set 
must be manipulated analogously to the training set:

```{r}
testdf1<-test[8:160]
testdf2<-testdf1[,-grep(pattern='avg',x=colnames(testdf1))]
testdf3<-testdf2[,-grep(pattern='max',x=colnames(testdf2))]
testdf4<-testdf3[,-grep(pattern='min',x=colnames(testdf3))]
testdf5<-testdf4[,-grep(pattern='stddev',x=colnames(testdf4))]
testdf6<-testdf5[,-grep(pattern='skewness',x=colnames(testdf5))]
testdf7<-testdf6[,-grep(pattern='kurtosis',x=colnames(testdf6))]
testdf8<-testdf7[,-grep(pattern='accel',x=colnames(testdf7))]
testdf9<-testdf8[,-grep(pattern='var',x=colnames(testdf8))]
testdf10<-testdf9[,-grep(pattern='amplitude',x=colnames(testdf9))]
```
Now that the test set is ready, a prediction can be made:

```{r}
pred<-predict(fit,newdata=testdf10)
```


